---
title: "Receptor-Based Drug Discovery"
author: "Emily Armbruster"
output: github_document
---
# We're going to look at the structure of the HIV protease and drug binding.
##### First, load the Bio3D package and use read.pdb() to pull in the structure with "library(bio3d)". Do this in the console so we don't re-load the package every time you knit the R Markdown file. Use "read.pdb" to bring in the "1hsg" pdb file. Then we're going to create two different pdb files for the ligand and the protein.

##### Assign the structure to the name "hiv". Get a quick summary of the object by typing in its name.

```{r}
library(bio3d)
hiv <- read.pdb("1hsg")
hiv
```

##### The two non protein resid values are "HOH (127)" and "MK1 (1)", representing water around the protein and the ligand.

###### Produce a "1hsg_protein.pdb" and "1hsg_ligand.pdb" file. Do the ligand first.
```{r}
ligand <- atom.select(hiv, "ligand", value = TRUE)
write.pdb(ligand, file = "1hsg_ligand.pdb")
```
##### Then do the protein.
```{r}
protein <- atom.select(hiv, "protein", value = TRUE)
write.pdb(protein, file = "1hsg_protein.pdb")
```

##### Format an output file with config.txt. 
##### out = all.pdbqt <- Type of file in output.

##### Inspect docking results. "multi=TRUE" allows us to read the file. View in VMD.
```{r}
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```

