---
title: "Class 06"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This is a H1

# Use hash-tag to make a header

This is my class 6 work.

## Use double hashtags to make a smaller heading

Use double astricks to make something **bold**

Insert some coding with the insert button above:

```{r}
plot(1:10)
```

You can tell it to insert different coding types with the insert button: python, R, bash, etc

Going to start parsing test files test1, test2, test3

```{r}
read.table("test1.txt", sep=",")
```
```{r}
read.table("test2.txt", sep="$")
```
```{r}
read.table("test3.txt", sep="\t")
```

#Functions! A function has 3 things: name (ei, read table), input argument (tells it what to do more specifically), body (actual work happens here). Note below, y is specified optionally.

```{r}
add <- function(x, y=1) {
  # Sum the input x and y 
  x + y}
```

#now use it...

```{r}
add(x=c(5, 6, 1, 100), y = 100)
```

#Each item separated by commma = different argument. This function only has two arguments... Adding more arguments throws an error.

#Can only take one data type. If you make an argument a different type of data, R throws an error

#If you don't give y a specific value, it will default to y.

```{r}
add(c(5, 5, NA, 7))
```

#use range function:

```{r}
x <- range (c(5, 2, 7, 10))
x[1] # min value
x[2] # max value
```

#test R rescale function

```{r}
rescale <- function(x) {
  rng <-range(x)
  (x - rng[1]) / (rng[2] - rng[1])
}
```

#Use it... If it fails, adjust, test again

```{r}
x <- c(1, 2, NA, 5)
range(x, na.rm=TRUE)
rescale2 <- function(x) {
  rng <- range(x)
  (x - rng[1]) / (rng[2] - rng[1])
}
```

#Take a look at the rng function!!

```{r}
rescale <- function(x, na.rm=TRUE, plot=FALSE) {
 if(na.rm) {
 rng <-range(x, na.rm=na.rm)
 } else {
 rng <-range(x)
 }
 print("Hello")
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 print("is it me you are looking for?")
 if(plot) {
 plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
}
```

```{r}
rescale(1:10)
```

#Handout Hands-on exercise:
#download package bio3d:
#install.packages("bio3d")
#Use function from package: issue library

```{r}
library(bio3d)
```

#What is the package? Has functions in it that are useful! including read.pdb: reads the crystal structure of protein...

```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor") 
```

#What kind of object is returned from the read.pdb() function?

```{r}
x <- read.pdb("4AKE")
is.list(x)
```

#What does the trim.pdb() function do?

```{r}
?trim.pdb
```

#What input parameter would turn off the marginal black and grey rectangles in the plots and what do they represent?

```{r}
plotb3(s3.b, typ="l", ylab="Bfactor") 
```
#The input parameter is "sse=s3.chainA". Omit by setting "sse=FALSE". They represent areas that form helices and sheets in the protein structure.

#What would be a better plot to compare across the different proteins? 

```{r}
x <- c(s1.b, s2.b, s3.b)

plotb3(x, sse=s1.chainA, typ="l", ylab="Bfactor")
```


